abxD01 Paper

### 2015-03-07 PAPER.Rmd
Edit first paragraph of Results and Fig 1 and S1 captions

***Levels of colonization resistance are associated with the initial structure of the gut microbiota.*** In mouse models of CDI, the baseline, untreated murine microbiota is completely resistant to *C. difficile* colonization. It was previously shown that C57Bl/6 mice treated with cefoperazone [], clindamycin [], or clindamycin in combination with a five antibiotic cocktail [] all have decreased resistance against this pathogen. To further characterize structural shifts in the microbiota associated with this loss of protective function, we chose a panel of seven representatives from six classes of antibiotics (**Table 1**) to differentially alter the microbiota and assess their resistance to *C. difficile* colonization. 16S rRNA gene sequencing of the fecal communities sampled at the time of *C. difficile* inoculation revealed distinct and reproducible microbiota structures (**Figure 1**). The community structures of the beta-lactam-treated mice (i.e. cefoperazone and ampicillin) were not significantly different from each other (AMOVA, p=0.-36). Clindamycin-treated microbiotas were significantly different from all other treatment groups (AMOVA, p=0.001) and were characterized by the predominance of *Enterobacteriaceae*. Of all groups significantly different from the untreated microbiota, the levels of *Enterobacteriaceae* increased with all treatments except with streptomycin. Vancomycin-treated communities saw the largest bloom in *Akkermansia*. In addition, the ciprofloxacin-treated microbiota were the only group that was not significantly different from untreated mice (AMOVA, p=0.09). Subsequently, 24 hours after inoculation *C. difficile* was undetectable in the untreated mice or the ciprofloxacin-treated mice; however it had successfully colonized the mice in all other treatment groups to similarly high levels (10<sup>7</sup>-10<sup>8</sup> CFU/g) (**Figure 1**). We calculated the diversity of all of the communities using the Inverse Simpson (**Supplemental Figure 1**) and Shannon indices (data not shown) and observed a significant negative correlation with subsequent *C. difficile* colonization (both p<0.001). These results suggest that multiple microbiota structures are capable of being colonized by *C. difficile* and that diversity may be an important factor in resistance to invasion.

@@

**Figure 1. Microbiota structures following diverse antibiotic treatments.** This figure shows the average percent relative abundances for family level phylotypes found in each experimental group on the day of *C. difficile* challenge. Each row depicts a different antibiotic perturbation. The level *C. difficile* colonization found 24 hours post microbiota sampling is plotted in the top right corner. Each represented phylum is a different shade of gray. Error bars depict standard error of the mean.

@@

**Supplemental Figure 1. Diversity is nevatively correlated with *C. difficile* levels.** This figure shows the relationship between the inverse simpson diversity index of microbiotas of each experimental group on day 0 and *C. difficile* CFU/g feces on day 1. Spearman's rho is plotted in the top right corner.

>>>>

***Levels of colonization resistance are associated with the initial structure of the gut microbiota.*** In mouse models of CDI, the baseline, untreated murine microbiota is completely resistant to *C. difficile* colonization. It was previously shown that C57Bl/6 mice treated with cefoperazone [], clindamycin [], or clindamycin in combination with a five antibiotic cocktail [] had decreased resistance against this pathogen. To further characterize structural shifts in the microbiota associated with this loss of protective function, we chose a panel of seven representatives from six classes of antibiotics (**Table 1**) to differentially alter the microbiota and assess their resistance to *C. difficile* colonization. 16S rRNA gene sequencing of the fecal communities sampled at the time of *C. difficile* inoculation revealed distinct and reproducible microbiota structures (**Figure 1**). The community structures of the beta-lactam-treated mice (i.e. cefoperazone and ampicillin) were not significantly different from each other (AMOVA, p=XXX0.36XXX). Clindamycin-treated microbiotas were significantly different from all other treatment groups (AMOVA, p=XXX0.001XXX) and were characterized by the predominance of *Enterobacteriaceae*. Of all groups significantly different from the untreated microbiota, the levels of *Enterobacteriaceae* increased with all treatments except with streptomycin. Vancomycin-treated communities saw the largest bloom in *Akkermansia*. In addition, the ciprofloxacin-treated microbiota were the only group that was not significantly different from untreated mice (AMOVA, XXXp=0.09XXX). Subsequently, 24 hours after inoculation *C. difficile* was undetectable in the untreated mice or the ciprofloxacin-treated mice; however it had successfully colonized the mice in all other treatment groups to similarly high levels (10<sup>7</sup>-10<sup>8</sup> CFU/g) (**Figure 1**). We calculated the diversity of all of the communities using the Inverse Simpson and Shannon indices and observed a significant negative correlation with subsequent *C. difficile* colonization (both **Supplemental Figure 1**; p<XXX0.001XXX). These results demonstrate that a variety of non-overlapping community structures can be sensitive to *C. difficile* colonization.

@@

**Figure 1. Microbiota structures following diverse antibiotic treatments.** This figure shows the average percent relative abundances for family level phylotypes found in each experimental group on the day of *C. difficile* challenge. Each row depicts a different antibiotic perturbation. The level *C. difficile* colonization found 24 hours post microbiota sampling is indicated in the top right for each treatment. Error bars indicate standard error of the mean.

@@

**Supplemental Figure 1. The inverse of Simpson's diversity index for the community on day 0 is negatively correlated with *C. difficile* colonization on day 1.** Spearman's rho is plotted in the top right corner.
